### 2020-07-14 paper.tex
simplify measure formalism for dictionaries

\textbf{Measure space for finite dictionaries.} We fix a countably infinite set $\mathcal{A}$ of possible keys, and for each key $a \in \mathcal{A}$,
we fix a measure space $(V_a, \Sigma_a, \mu_a)$ determining the possible values $V_a$ that may be
associated with $a$. Given a set of keys $A$, we can consider the set $V_A = \times_{a \in A} V_a$ of \textit{assignments} of values to each key. ($V_A$ can also be made a measure space, using the standard product $\sigma$-algebra $\Sigma_A = \otimes_{a \in A} \Sigma_a$ and the product measure $\mu_A = \times_{a \in A} \mu_a$.)  Then the set of all finite dictionaries is $\mathcal{D} := \bigcup_{A \subset \mathcal{A}, |A| < \infty} \{(A, \mathbf{x}) \mid \mathbf{x} \in V_A\}$. That is, a dictionary specifies 
a finite set of keys $A \subset \mathcal{A}$ at which it has values, and an assignment $\mathbf{x}$ of values $x_a$ for each.
We equip $\mathcal{D}$ with the $\sigma$-algebra $\Sigma_{\mathcal{D}} := \{\bigcup_{A \subset \mathcal{A}, |A| < \infty} \{(A, \mathbf{x}) \mid \mathbf{x} \in B_A\} \mid B_A \in \Sigma_A \, \text{for each finite } A \subset \mathcal{A}\}$ to obtain a measurable space
of dictionaries. A reference measure $\mu_\mathcal{D}$ on this space can be constructed using the product measures $\mu_A$: we set $\mu(B) := \sum_{A \in \mathcal{A}, |A| < \infty} \mu_A(\{\mathbf{x} \mid (A, \mathbf{x}) \in B\})$.

\textbf{Notation for dictionaries.} Given a dictionary $x = (A, \mathbf{x})$, we write $A_m$ for $A$ and $m[a]$ for the value $x_a$ associated with a key $a \in A$.

\textbf{Probability distributions on finite dictionaries.} 
When all keys are discrete (have countable $V_a$ and counting measure for $\mu_{a}$), a probability distribution on dictionaries is defined by a $p$ is a probability mass function $p : \mathcal{D} \to [0, 1]$ that assigns a probability $p(m)$ to each dictionary $m \in \mathcal{P}$ such that $\sum_{x \in \mathcal{D}} p(m) = 1$.
More generally a probability distribution in dictionaries is defined by a probability density $p : \mathcal{D} \rightarrow [0, \infty)$ such that $\int p(m) \mu_\mathcal{D}(\text{d}m) = 1$. We now give several examples
to build intuition.

@@

Suppose that the model distribution $p$ and auxiliary distributions $q_x$ are all probability distributions on dictionaries.
Then, $X$ and $Y$ are both sets of dictionaries, and the joint space $Z$ is a set of pairs $(x, y)$ of dictionaries.
Suppose that both $X$ and $Y$ use the set of keys $\mathcal{A}$, and that $\mathcal{A}$ can be partitioned $\mathcal{A} = \mathcal{I} \cup \mathcal{K}$ where is a set of \emph{discrete keys} $\mathcal{I}$ that use countable $V_a$ and counting measure $\mu_a$ for $a \in \mathcal{I}$ and $\mathcal{K}$ is a set of \emph{continuous keys} that use $V_a = \mathbb{R}^{d_a}$ for some $d_a$ and Lebesgue-measure $\mu_a$ for $a \in \mathcal{K}$.
%And suppose that there exists an involution $f : Z \to Z$ such that (i) 
Suppose there is a countable partition of $Z$ into $\{Z_e : e \in E\}$ such that across each $Z_e$ the sets of keys $A_x$ and $A_y$ and the values $\mathbf{x}_{\mathrm{d}}$ and $\mathbf{y}_{\mathrm{d}}$ of discrete keys are both constant (so that each set $Z_e$ is isomorphic to a Euclidean space of assignments to the continuous keys in the two dictionaries).
Suppose there is an involution $g : E \to E$ between elements of the partition (a function with $g = g^{-1}$), and a family of continuously differentiable bijections $h_{e} : Z_e \to Z_{g(e)}$ for each $e$ with $h_{e} = h_{g(e)}^{-1}$.
Let $e(z) \in E$ denote the element of the partition for an alement $z \in Z$.
Then, $f : Z \to Z$ given by $f(z) := h_{e(z)}(z)$ is an involution:
\begin{equation}
f(f(z)) = h_{e(f(z))}(h_{e(z)}(z)) = h_{g(e(z))}(h_{e(z)}(z)) = z
\end{equation}
Let $|J h_e|(z)$ denote the absolute value of the Jacobian (determinant) of $h_e$, evaluated at $z$.
Algorithm~\ref{alg:involutive-mcmc} defines involutive MCMC in this more structured setting, in which the Radon-Nikodym is replaced with a Jacobian (see Section~\ref{sec:radon-nikodym-special-case} for a derivation).
\begin{algorithm}[h]
\begin{algorithmic}
\Procedure{involutive-mcmc-kernel}{$p$, $q$, $f$, $x$}
    \State $y \sim q_x(\cdot)$ \Comment{Sample auxiliary dictionary}
    \State $(x', y') \gets f(x, y)$ \Comment{Apply involution}
    \State $\alpha \gets
        \displaystyle \frac{p(x') q_{x'}(y')}{p(x) q_{x}(y)} \cdot |J h_e|(z)$
    \State $r \sim \mathrm{Uniform}(0, 1)$
    \State \algorithmicif \, $r \le \alpha$ \algorithmicthen \, \Return $x'$ \algorithmicelse \, \Return $x$ 
\EndProcedure
\end{algorithmic}
\caption{Involutive MCMC for Dictionaries}
\label{alg:involutive-mcmc}
\end{algorithm}

>>>>

\textbf{The space of finite dictionaries.}
We fix a countably infinite set $\mathcal{A}$ of possible keys, such that each key $a$ is either \emph{discrete} ($a \in \mathcal{I}$) or \emph{continuous} ($a \in \mathcal{J}$), where $\mathcal{A} = \mathcal{I} \cup \mathcal{J}$.\footnote{
It is straighforward to assign a general meaure space to each address, but this is not necessary for our purposes.}
Let $V_a$ denote the set of possible values for key $a$, where $V_a$ is a countable set for each discrete key, and where $V_a = \mathbb{R}^{d_a}$ for each continuous key for some $d_a$.
Given a set of keys $A$, let $V_A = \times_{a \in A} V_a$ denote the set of \textit{assignments} of values to each key.
Then the set of all finite dictionaries is $\mathcal{D} := \bigcup_{A \subset \mathcal{A}, |A| < \infty} \{(A, \mathbf{x}) \mid \mathbf{x} \in V_A\}$. That is, a dictionary specifies 
a finite set of keys $A \subset \mathcal{A}$ at which it has values, and an assignment $\mathbf{x}$ of values $x_a$ for each.

\textbf{A measure space of finite dictionaries.}
We associate a measure $\mu_a$ on $V_a$ for each key---the counting measure for each discrete key and the Lebesgue-measure on $\mathbf{R}^{d_a}$ for each continuous key.
For each finite set of keys $A$, we make $V_A$ a measure space using the standard product $\sigma$-algebra $\Sigma_A = \otimes_{a \in A} \Sigma_a$ and the product measure $\mu_A = \times_{a \in A} \mu_a$.
We equip $\mathcal{D}$ with the $\sigma$-algebra $\Sigma_{\mathcal{D}} := \{\bigcup_{A \subset \mathcal{A}, |A| < \infty} \{(A, \mathbf{x}) \mid \mathbf{x} \in B_A\} \mid B_A \in \Sigma_A \, \text{for each finite } A \subset \mathcal{A}\}$ to obtain a measurable space
of dictionaries.
A reference measure $\mu_\mathcal{D}$ on this space can be constructed using the product measures $\mu_A$: we set $\mu(B) := \sum_{A \in \mathcal{A}, |A| < \infty} \mu_A(\{\mathbf{x} \mid (A, \mathbf{x}) \in B\})$.

\textbf{Notation for dictionaries.} Given a dictionary $x = (A, \mathbf{x})$, we write $A_m$ for $A$ and $m[a]$ for the value $x_a$ associated with a key $a \in A$.
We also denote specific dictionaries using notation $\{k_1 \mapsto v_1, k_2 \mapsto v_2, \ldots\}$.
For example, the dictionary $(A, \mathbf{x})$ with $A = \{1, \mathtt{"foo"}\}$ and $x_1 = 0.123$ and $x_{\mathtt{"foo"}} = 5$ is denoted 
$\{1 \mapsto 0.123, \mathtt{"foo"} \mapsto 5\}$.

\textbf{Probability distributions on finite dictionaries.} 
When all keys are discrete, a probability distribution on dictionaries is defined by a $p$ is a probability mass function $p : \mathcal{D} \to [0, 1]$ that assigns a probability $p(m)$ to each dictionary $m \in \mathcal{P}$ such that $\sum_{x \in \mathcal{D}} p(m) = 1$.
More generally a probability distribution in dictionaries is defined by a probability density $p : \mathcal{D} \rightarrow [0, \infty)$ such that $\int p(m) \mu_\mathcal{D}(\text{d}m) = 1$. We now give several examples
to build intuition.

@@

Suppose that the model distribution $p$ and auxiliary distributions $q_x$ are all probability distributions on dictionaries.
Then, $X$ and $Y$ are both sets of dictionaries, and the joint space $Z$ is a set of pairs $(x, y)$ of dictionaries with keys $\mathcal{A}$, so that $Z = X \times Y \subseteq \mathcal{D} \times \mathcal{D}$.
%And suppose that there exists an involution $f : Z \to Z$ such that (i) 
Suppose there is a countable partition of $Z$ into $\{Z_e : e \in E\}$ such that across each $Z_e$ the sets of keys $A_x$ and $A_y$ and the values $x_a$ and $y_a$ for all discrete keys $a$ are both constant (so that each set $Z_e$ is isomorphic to a Euclidean space of assignments to the continuous keys in the two dictionaries).
Suppose there is an involution $g : E \to E$ between elements of the partition (a function with $g = g^{-1}$), and a family of continuously differentiable bijections $h_{e} : Z_e \to Z_{g(e)}$ for each $e$ with $h_{e} = h_{g(e)}^{-1}$.
Let $e(z) \in E$ denote the element of the partition for an alement $z \in Z$.
Then, $f : Z \to Z$ given by $f(z) := h_{e(z)}(z)$ is an involution:
\begin{equation}
f(f(z)) = h_{e(f(z))}(h_{e(z)}(z)) = h_{g(e(z))}(h_{e(z)}(z)) = z
\end{equation}
Let $|J h_e|(z)$ denote the absolute value of the Jacobian (determinant) of $h_e$, evaluated at $z$.
Algorithm~\ref{alg:involutive-mcmc} defines involutive MCMC in this more structured setting, in which the Radon-Nikodym is replaced with a Jacobian (see Section~\ref{sec:radon-nikodym-special-case} for a derivation).
\begin{algorithm}[h]
\begin{algorithmic}
\Procedure{involutive-mcmc-kernel}{$p$, $q$, $f$, $x$}
    \State $y \sim q_x(\cdot)$ \Comment{Sample auxiliary dictionary}
    \State $(x', y') \gets f(x, y)$ \Comment{Apply involution}
    \State $\alpha \gets
        \displaystyle \frac{p(x') q_{x'}(y')}{p(x) q_{x}(y)} \cdot |J h_e|(z)$
    \State $r \sim \mathrm{Uniform}(0, 1)$
    \State \algorithmicif \, $r \le \alpha$ \algorithmicthen \, \Return $x'$ \algorithmicelse \, \Return $x$ 
\EndProcedure
\end{algorithmic}
\caption{Involutive MCMC for Dictionaries}
\label{alg:involutive-mcmc}
\end{algorithm}
